Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-23 thesis/paper_draft.qmd
update appendix, include codes for sim supplemetary, reproducibility

<!-- ![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-results-2} -->

```{r}
#| label: fig-sim-results-2
#| fig-cap: "Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base."
#| fig.height: 10
#| fig.width: 10
#| out.width: "100%"

@@

energy_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
  energy_plots[[2]] + labs(subtitle = "T=300") +
  energy_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
  energy_plots[[4]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank())

```

Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score, a proper scoring rule for multivariate predictive distributions. @fig-sim-results-2 presents the percentage relative improvement in energy score for 1-step-ahead forecasts. Across both hierarchies and time dimensions, the MinT methods consistently outperform the base forecasts. Meanwhile, the differences among *MinT-S* and *MinT-N* variants are small, except for the 6 by 6 case with 50 observations, where shrinkage has a pronounced edge. The PC-adjusted variants again degrade performance as we add more principal components. The scaled variance and $h$-step-ahead residuals approaches are not available since we only evaluate 1-step-ahead covariance estimates.

Other scoring rules, such as the Winkler score and CRPS, yield similar conclusions.

## Other Data Generating Processes

In attempts to differentiate the performance of NOVELIST from the shrinkage estimator, we also simulate from a sparse covariance matrix for the previous section's 2 groups of 50 bottom series setting, as illustrated in @fig-Sigma-2x50. The sparse covariance matrix is obtained by randomly choosing 40% of the bottom series and setting their correlations with all other series to zero, resulting in a grid-like sparse structure. The VAR(1) coefficient matrix remains the same as in the dense case. The idea is to allow NOVELIST to exploit the sparsity in the covariance structure, since it can control the sparsity of the shrinkage target via the thresholding parameter $\delta$. However, no profound insights can be drawn from the results.

@@

```{r summary structure tourism}
#| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"
#| label: tbl-tourism-structure

@@

All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov].

@@

@@

## Appendix: Australian Domestic Tourism Geographical Hierarchy {#sec-tourism-supplement}

@@

>>>>

<!-- ![Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base.](D:/Github/Recon_Honours_Thesis/figs/sim_energy.png){#fig-sim-energy} -->

```{r}
#| label: fig-sim-energy
#| fig-cap: "Percentage relative improvement in Energy score in both hierarchies and time dimensions, for 1-step-ahead forecasts. The positive entries indicate a decrease in Energy score relative to base."
#| fig.height: 10
#| fig.width: 10
#| out.width: "100%"

@@

energy_plots[[1]] + labs(title = "6 Groups of 6", subtitle = "T=50") +
  energy_plots[[2]] + labs(subtitle = "T=300") +
  energy_plots[[3]] + labs(title = "2 Groups of 50", subtitle = "T=50") +
  energy_plots[[4]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank())
```

Moving on to probabilistic forecasts, we evaluate the performance of reconciliation methods using the energy score, a proper scoring rule for multivariate predictive distributions. @fig-sim-energy presents the percentage relative improvement in energy score for 1-step-ahead forecasts. Across both hierarchies and time dimensions, the MinT methods consistently outperform the base forecasts. Meanwhile, the differences among *MinT-S* and *MinT-N* variants are small, except for the 6 by 6 case with 50 observations, where shrinkage has a pronounced edge. The PC-adjusted variants again degrade performance as we add more principal components. The scaled variance and $h$-step-ahead residuals approaches are not available since we only evaluate 1-step-ahead covariance estimates.

Other scoring rules, such as the Winkler score and CRPS, yield similar conclusions. More detailed results are provided in Appendix @sec-sim-supplement.

## Other Data Generating Processes {#sec-sim-other}

In attempts to differentiate the performance of NOVELIST from the shrinkage estimator, we also simulate from a sparse covariance matrix for the previous section's 2 groups of 50 bottom series setting, as illustrated in @fig-Sigma-2x50. The sparse covariance matrix is obtained by randomly choosing 40% of the bottom series and setting their correlations with all other series to zero, resulting in a grid-like sparse structure. The VAR(1) coefficient matrix remains the same as in the dense case. The idea is to allow NOVELIST to exploit the sparsity in the covariance structure, since it can control the sparsity of the shrinkage target via the thresholding parameter $\delta$. However, no profound insights can be drawn from the results (provided in Appendix @sec-sim-supplement).

@@

```{r}
#| label: tbl-tourism-structure
#| tbl-cap: "Hierarchical and grouped structure of Australian domestic tourism flows"

@@

All data generation, covariance estimation, and reconciliation routines were implemented in the ReconCov R package and is available under an open‐source license on GitHub [@ReconCov]. Regarding reproducibility, see Appendix @sec-reproducibility.

@@

\vspace{-0.5em}

@@

\vspace{-0.5em}

**Supplementary figures for simulation study in @sec-sim-hierarchy-size.**

```{r}
#| label: fig-appendix-sim-winkler
#| fig-cap: "Percentage relative improvement in the Winkler score at 80% (left column) and 95% (right column) nominal coverage of different 1-step-ahead reconciled forecasts over the base forecasts for both hierarchical structures and time series lengths. The positive entries indicate a decrease in Winkler score relative to base."
#| fig.height: 10
#| fig.width: 8.5
#| out.width: "83%"
#| fig-pos: "H"
sim_plots <- list()

for (sim_name in names(sim_prob)) {
  for (rule_name in c("W80", "W95")) {
    p <- sim_prob[[sim_name]] |>
      filter(rule == rule_name) |>
      filter(model_label != "Base") |>
      ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
        geom_col(color = "black") +
        scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
        recon_fill_scale +
        labs(x = NULL, y = "% improvements") +
        theme_minimal() +
        theme(
          axis.text.x = element_text(angle = 20, hjust = 0.5)
        )
    # remove x axis labels for all but last row
    if (sim_name != "S100_T300") p <- p + theme(axis.text.x = element_blank())
    name <- paste0(sim_name, "_", rule_name)
    sim_plots[[name]] <- p
  }
}

sim_plots[[1]] + labs(title = "Winkler score 80%", subtitle = "6 Groups of 6, T=50") +
  sim_plots[[2]] + labs(title = "Winkler score 95%", subtitle = "6 Groups of 6, T=50") +
  sim_plots[[3]] + labs(subtitle = "6 Groups of 6, T=300") +
  sim_plots[[4]] + labs(subtitle = "6 Groups of 6, T=300") +
  sim_plots[[5]] + labs(subtitle = "2 Groups of 50, T=50") +
  sim_plots[[6]] + labs(subtitle = "2 Groups of 50, T=50") +
  sim_plots[[7]] + labs(subtitle = "2 Groups of 50, T=300") +
  sim_plots[[8]] + labs(subtitle = "2 Groups of 50, T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "right", legend.title = element_blank())
```

To complement the main simulation probabilistic forecast results (using Energy score) in @sec-sim-hierarchy-size, @fig-appendix-sim-winkler presents percentage relative improvements in Winkler scores at 80% and 95% nominal coverage for 1-step-ahead forecasts across both hierarchical structures (small and large) and time series lengths (short and long). The patterns largely mirror those observed with the Energy score in @fig-sim-energy.

\vspace{3em}

**Supplementary figure for simulation study with sparse covariance structure in @sec-sim-other.**

```{r}
#| label: fig-appendix-sim-sparse-MSE
#| fig-cap: "Percentage relative improvement in MSE of 1-to-12-step reconciled forecasts for sparse covariance (of the innovation process) setting, for the large 2-groups-of-50 hierarchy and both time series lengths. The positive entries indicate a decrease in MSE."
#| fig.height: 6
#| fig.width: 10
#| out.width: "100%"
#| fig-pos: "H"
sim_sparse_MSE <- readRDS(paste0(path_lib, "sim_sparse_MSE.rds"))

sparse_MSE_plots <- list()

cap <- -6
for (sim_name in names(sim_sparse_MSE)) {
  plot_df <- sim_sparse_MSE[[sim_name]] |> 
    mutate(
      capped = pct_change <= cap,
      capped_pct_change = pmax(pct_change, cap)
    ) |> 
    mutate(shape_lab = ifelse(shape_lab %in% c("PC1","PC2"), shape_lab, ""))
  p <- plot_df |>
    ggplot(aes(
      x = h, y = capped_pct_change,
      color = family, shape = shape_lab, linetype = variant_type
    )) +
      geom_line(linewidth = 0.8) +
      geom_point(alpha = 0.85, size = 2.6, na.rm = TRUE) +
      recon_scales +
      theme_minimal() +
      labs(x = "Forecast Horizon", y = "% improvements")
  sparse_MSE_plots[[sim_name]] <- p
}

sparse_MSE_plots[[1]] + labs(subtitle = "T=50") +
  sparse_MSE_plots[[2]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.direction = "vertical", legend.title.position = "left")
```

```{r}
#| label: fig-appendix-sim-sparse-energy
#| fig-cap: "Percentage relative improvement in Energy score of 1-step reconciled forecasts for sparse covariance (of the innovation process) setting, for the large 2-groups-of-50 hierarchy and both time series lengths. The positive entries indicate a decrease in Energy score."
#| fig.height: 6
#| fig.width: 10
#| out.width: "90%"
#| fig-pos: "H"
sim_sparse_prob <- readRDS(paste0(path_lib, "sim_sparse_prob.rds"))
sparse_energy_plots <- list()

for (sim_name in names(sim_sparse_prob)) {
  p <- sim_sparse_prob[[sim_name]] |>
    filter(rule == "Energy") |>
    filter(model_label != "Base") |>
    ggplot(aes(x = model_label, y = pct_change_avg, fill = family, alpha=pc_k)) +
      geom_col(color = "black") +
      scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
      recon_fill_scale +
      labs(x = "Model", y = "% improvements") +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 20, hjust = 0.5)
      )
  sparse_energy_plots[[sim_name]] <- p
}

sparse_energy_plots[[1]] + labs(subtitle = "T=50") +
  sparse_energy_plots[[2]] + labs(subtitle = "T=300") +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank())
```

\vspace{1em}

## Appendix: Description of Australian Geographical Hierarchy {#sec-tourism-supplement}

@@

## Appendix: Reproducibility {#sec-reproducibility}

This research project has 3 Github repositories:

1. **ReconCov**: The main *R package* implements the studied covariance estimators, MinT reconciliation, general data generating process functions. Refer to [@ReconCov] for installation and documentation.

2. **Recon_Honours_Analysis**: This repository contains all R codes to produce the simulation studies and empirical analysis in this thesis (including scripts to run on a high-performance cluster), along with wrangling, processing, and analysing the results data. It depends on the *ReconCov* package. The access is upon request.

3. **Recon_Honours_Thesis**: This repository contains the thesis manuscript in Quarto format, along with all figures and tables. The access is upon request.

Please contact the author via email `Vincent.Su@monash.edu` for access to the analysis and thesis repositories.
